% First parameter can be changed eg to "Glossary" or something.
% Second parameter is the max length of bold terms.
%To keep this glossary useful, we mainly focus on abbreviations specific to our domain and don't list very common abbrevations (such as NLP - Natural Language Processing).
\begin{mclistof}{Glossary}{4.8cm}
%\item[NLP] Natural Language Processing
%\item[CV] Computer Vision
\item[Splicing] Conversion of pre-mRNA to mature mRNA through removal of introns and joining together of exons.
\item[Exon] Genomic sequence which is kept from the pre-mRNA during splicing.
\item[Intron] Genomic sequence which is removed from the pre-mRNA during splicing.
\item[Non-coding] Genomic sequence which is not directly used during protein construction. 
%Process by which a pre-mRNA is converted to a mature mRNA or the actual act of cutting out genomic sequences in that process itself.
\item[Junction] Site in a mature mRNA transcript where two exons border each other.
\item[Cis-regulatory element] Non-coding DNA region on the same DNA molecule as the gene (`close to the gene') it regulates.
%\item[Sequence information] Information contained in a sequence of nucleotides of the DNA.
\item[PSI] Percent-spliced in, the proportion of transcripts containing a specific exon (or junction).
\item[Motif] A widespread nucleotide sequence pattern conjectured to be biologically significant.
%\item[MLP] Multi-layer perceptron
%\item[w2v] word2vec algorithm
%\item[D2V] Either the algorithm document2vector, or the specific model based on this algorithm.
\item[EST] Expressed sequence tag, a short cDNA sequence generated through Sanger sequencing.
\item[GTEx] Genotype-Tissue Expression, a project providing a large repository of human-sequencing data.

\item[iPSC] induced pluripotent stem cells, mature cells which have been reprogrammed to an immature pluripotent (undetermined) state.
\item[HipSci] Human Induced Pluripotent Stem Cell Initiative, an initiative proving a large repository of iPSC-based sequencing data.

%\item[ROC] Receiver Operating Curve
%\item[AUC] Area under ROC curve
%\item[Seq2seq] Sequence-to-sequence learning, a framework often used in MT.
%\item[MT] Machine Translation
%\item[RNN] Recurrent Neural Network
%\item[CNN] Convolutional Neural Network
\item[Splicing Code] A set of rules describing the regulatory mechanisms underlying splicing. Often implemented as a computational model.
\item[DSC] Deep Splicing Code, a model used for constitutive exon classification.



\end{mclistof} 
